---
title: "oneR: Automated Statistical Testing with Normality Assessment"
author: "Bright Boamah"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

The oneR package provides a streamlined approach to statistical testing by automatically assessing data normality and selecting appropriate parametric or non-parametric tests. This vignette demonstrates the package's functionality and provides comprehensive examples for various use cases.

## Package Philosophy

Statistical testing often requires researchers to make decisions about which test to use based on data characteristics, particularly normality assumptions. The oneR package automates this decision-making process while maintaining statistical rigor and providing comprehensive documentation of the analysis process.

## Key Features

- **Automated test selection**: Chooses between parametric and non-parametric tests based on normality assessment
- **Comprehensive visualization**: Generates plots for normality assessment and test results
- **PDF reporting**: Creates detailed reports with all analysis components
- **Flexible interface**: Supports one-sample, two-sample, and paired tests
- **Statistical rigor**: Uses established tests (Shapiro-Wilk, t-test, Wilcoxon) with proper implementation

# Installation and Setup

```{r setup, eval=FALSE}
# Install required packages if not already installed
install.packages(c("ggplot2", "gridExtra", "knitr", "rmarkdown"))

# Load the oneR package
library(oneR)
```

```{r load-package, echo=FALSE, message=FALSE}
# For vignette building, we'll simulate loading the package
library(stats)
library(graphics)
library(ggplot2)
library(gridExtra)
library(knitr)
```

# Basic Usage

## One-Sample Testing

The most basic use case involves testing whether a single sample comes from a population with a specific mean.

```{r one-sample-example, eval=FALSE}
# Generate sample data
set.seed(123)
sample_data <- rnorm(30, mean = 100, sd = 15)

# Perform automated testing
result <- oneR_test(sample_data, mu = 100)

# View results
print(result)
```

### Understanding the Output

The oneR_test function returns a comprehensive object containing:

- **Normality assessment**: Results from Shapiro-Wilk tests
- **Test selection**: Which statistical test was chosen and why
- **Test results**: Complete statistical output including p-values and confidence intervals
- **Recommendations**: Interpretation of results in plain language

## Two-Sample Testing

Comparing two independent groups is equally straightforward:

```{r two-sample-example, eval=FALSE}
# Generate two groups
set.seed(456)
group_a <- rnorm(25, mean = 85, sd = 12)
group_b <- rnorm(25, mean = 90, sd = 12)

# Perform comparison
comparison <- oneR_test(group_a, group_b)

# View summary
summary(comparison)
```

## Paired Testing

For paired or repeated measures data:

```{r paired-example, eval=FALSE}
# Generate paired data
set.seed(789)
before <- rnorm(20, mean = 75, sd = 10)
after <- before + rnorm(20, mean = 5, sd = 3)  # Treatment effect

# Perform paired test
paired_result <- oneR_test(before, after, paired = TRUE)
print(paired_result)
```

# Advanced Features

## Customizing Analysis Parameters

The oneR_test function provides several parameters for customizing the analysis:

```{r advanced-parameters, eval=FALSE}
# Custom significance levels and alternatives
result <- oneR_test(
  x = sample_data,
  mu = 95,
  alternative = "greater",  # One-tailed test
  alpha = 0.01,            # Stricter significance level
  conf.level = 0.99        # Higher confidence level
)
result
```

### Parameter Guide

| Parameter | Description | Default | Options |
|-----------|-------------|---------|---------|
| `x` | Primary data vector | Required | Numeric vector |
| `y` | Second group (optional) | NULL | Numeric vector |
| `mu` | Hypothesized mean | 0 | Numeric value |
| `alternative` | Alternative hypothesis | "two.sided" | "two.sided", "less", "greater" |
| `alpha` | Significance level for normality | 0.05 | 0 < α < 1 |
| `conf.level` | Confidence level for test | 0.95 | 0 < level < 1 |
| `paired` | Paired test indicator | FALSE | TRUE/FALSE |

# Visualization Capabilities

## Normality Assessment Plots

The package generates comprehensive plots to assess data normality:

```{r normality-plots, eval=FALSE}
# Generate normality assessment plots
plot_normality(result)
```

These plots include:

- **Histograms** with normal distribution overlay
- **Q-Q plots** with reference lines
- **Boxplots** with Shapiro-Wilk p-values

## Test Results Visualization

Results can be visualized to understand the data and test outcomes:

```{r results-plots, eval=FALSE}
# Generate test results plots
plot_results(comparison)
```

For two-sample tests, this includes:

- **Side-by-side boxplots** for group comparison
- **Density plots** showing distribution overlap
- **Summary panels** with key statistics

## Integrated Plotting

The plot method provides convenient access to all visualizations:

```{r integrated-plots, eval=FALSE}
# Show all plots
plot(result, type = "both")

# Show only normality plots
plot(result, type = "normality")

# Show only results plots
plot(result, type = "results")
```

# PDF Report Generation

## Basic Report Generation

Creating comprehensive PDF reports is straightforward:

```{r basic-report, eval=FALSE}
# Generate standard report
library(latexpdf)
oneR_report_pdf(result, "analysis_report.pdf")
```

## Customized Reports

Reports can be customized with various options:

```{r custom-report, eval=FALSE}
# Customized report with raw data
oneR_report(
  result,
  output_file = "detailed_analysis.pdf",
  title = "Clinical Trial Statistical Analysis",
  author = "Research Team",
  include_data = TRUE
)
```

### Report Contents

Generated reports include:

1. **Executive Summary**: Key findings and conclusions
2. **Data Overview**: Descriptive statistics for all groups
3. **Normality Assessment**: Detailed Shapiro-Wilk test results
4. **Statistical Test Results**: Complete test output with interpretation
5. **Visualizations**: All relevant plots embedded in the document
6. **Conclusions and Recommendations**: Statistical interpretation and guidance
7. **Technical Details**: Software information and methodology
8. **Raw Data** (optional): Complete dataset tables

# Real-World Examples

## Example 1: Clinical Trial Analysis

```{r clinical-example, eval=FALSE}
# Simulate clinical trial data
set.seed(2024)
placebo_group <- rnorm(50, mean = 120, sd = 15)  # Blood pressure
treatment_group <- rnorm(50, mean = 110, sd = 15)

# Perform analysis
clinical_analysis <- oneR_test(placebo_group, treatment_group)

# Generate comprehensive report
oneR_report(
  clinical_analysis,
  "clinical_trial_report.pdf",
  title = "Blood Pressure Treatment Efficacy Analysis",
  author = "Clinical Research Team"
)
```

## Example 2: Quality Control Testing

```{r qc-example, eval=FALSE}
# Manufacturing quality control data
set.seed(2024)
production_measurements <- c(
  rnorm(30, mean = 50.2, sd = 0.8),  # Normal production
  runif(5, min = 48, max = 52)       # Some outliers
)

# Test against specification
qc_result <- oneR_test(
  production_measurements, 
  mu = 50.0,
  alternative = "two.sided"
)

# Quick assessment
print(qc_result)
```

## Example 3: Educational Assessment

```{r education-example, eval=FALSE}
# Pre/post training scores
set.seed(2024)
pre_scores <- rbeta(25, 2, 3) * 100   # Skewed distribution
post_scores <- pre_scores + rnorm(25, mean = 8, sd = 5)

# Paired analysis
training_effect <- oneR_test(pre_scores, post_scores, paired = TRUE)

# Visualization
plot(training_effect, type = "both")
```

# Statistical Methodology

## Normality Testing

The package uses the Shapiro-Wilk test for normality assessment, which is appropriate for sample sizes up to 5000. The test evaluates the null hypothesis that the data comes from a normal distribution.

### Test Selection Logic

The automated test selection follows this decision tree:

1. **Assess normality** for all groups using Shapiro-Wilk test
2. **If all groups are normal** (p > α):
   - Use parametric tests (t-tests)
3. **If any group is non-normal** (p ≤ α):
   - Use non-parametric tests (Wilcoxon tests)

### Parametric Tests Used

- **One-sample t-test**: Compares sample mean to hypothesized value
- **Two-sample t-test**: Compares means of two independent groups
- **Paired t-test**: Compares paired observations

### Non-Parametric Tests Used

- **Wilcoxon signed-rank test**: Non-parametric alternative to one-sample and paired t-tests
- **Wilcoxon rank-sum test**: Non-parametric alternative to two-sample t-test

## Advantages of Automated Selection

1. **Reduces user error** in test selection
2. **Ensures appropriate assumptions** are met
3. **Provides consistent methodology** across analyses
4. **Documents decision rationale** in reports

# Best Practices

## Data Preparation

Before using oneR, ensure your data is properly prepared:

```{r data-prep, eval=FALSE}
# Remove missing values
clean_data <- data[!is.na(data)]

# Check for outliers
boxplot(clean_data)

# Verify data types
str(clean_data)
```

## Sample Size Considerations

- **Minimum sample size**: 3 observations (required for Shapiro-Wilk test)
- **Recommended minimum**: 10-15 observations for reliable normality assessment
- **Maximum for Shapiro-Wilk**: 5000 observations

## Interpretation Guidelines

### P-value Interpretation

- **p < α**: Reject null hypothesis (statistically significant)
- **p ≥ α**: Fail to reject null hypothesis (not statistically significant)

### Effect Size Considerations

While oneR focuses on statistical significance, consider practical significance:

- **Confidence intervals**: Indicate range of plausible effect sizes
- **Mean differences**: Assess practical importance
- **Context**: Consider domain-specific meaningful differences

## Common Pitfalls to Avoid

1. **Multiple comparisons**: Adjust significance levels when performing multiple tests
2. **Sample size**: Ensure adequate power for detecting meaningful effects
3. **Assumptions**: Verify that chosen tests are appropriate for your data
4. **Interpretation**: Distinguish between statistical and practical significance

# Troubleshooting

## Common Issues and Solutions

### Issue: "Data must have at least 3 observations"

**Solution**: Ensure your data vectors have sufficient observations for normality testing.

```{r troubleshoot-1, eval=FALSE}
# Check data length
length(your_data)

# Remove missing values
clean_data <- your_data[!is.na(your_data)]
```

### Issue: PDF generation fails

**Solution**: The package will automatically generate HTML reports if PDF creation fails.

```{r troubleshoot-2, eval=FALSE}
# Check if required packages are installed
install.packages(c("rmarkdown", "knitr"))

# Try HTML output explicitly
oneR_report(result, "report.html")
```

### Issue: Plots not displaying correctly

**Solution**: Ensure graphics devices are properly configured.

```{r troubleshoot-3, eval=FALSE}
# Reset graphics parameters
dev.off()

# Try different plot types
plot(result, type = "normality")
```

# Extending oneR

## Custom Plotting

You can create custom visualizations using the data from oneR objects:

```{r custom-plots, eval=FALSE}
# Extract data from oneR result
x_data <- result$data_x
y_data <- result$data_y

# Create custom plots
library(ggplot2)
custom_plot <- ggplot(data.frame(x = x_data), aes(x = x)) +
  geom_histogram(bins = 20, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Custom Data Visualization")

print(custom_plot)
```

## Integration with Other Packages

oneR results can be easily integrated with other statistical packages:

```{r integration, eval=FALSE}
# Extract results for further analysis
test_results <- extract_results(result)

# Use with other packages
library(broom)
tidy_results <- broom::tidy(result$test_result)
```

# Conclusion

The oneR package provides a comprehensive solution for automated statistical testing with proper normality assessment. By combining rigorous statistical methodology with user-friendly interfaces and comprehensive reporting, oneR enables researchers to conduct reliable statistical analyses efficiently.

Key benefits include:

- **Automated decision-making** reduces errors in test selection
- **Comprehensive visualization** aids in data understanding
- **Professional reporting** facilitates communication of results
- **Statistical rigor** ensures appropriate methodology

For additional support and examples, consult the package documentation and help files.

# References

1. Shapiro, S. S., & Wilk, M. B. (1965). An analysis of variance test for normality (complete samples). *Biometrika*, 52(3/4), 591-611.

2. Wilcoxon, F. (1945). Individual comparisons by ranking methods. *Biometrics Bulletin*, 1(6), 80-83.

3. Student. (1908). The probable error of a mean. *Biometrika*, 6(1), 1-25.

4. R Core Team (2024). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria.

